---
title: "583 Project Report"
output: pdf_document
---


```{r, include = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(gridExtra)
library(glmnet)
library(np)
library(nortest)
library(MASS)
library(glmnet)
library(leaps)
library(grid)
library(cowplot)
library(corrplot)
library(knitr)
library(mgcv)
library(randomForest)
library(gbm)
#library(rgl)

#setwd('DATA 583/583-project')
raw_data <- read.csv("raw_data.csv")

# Create new variables
raw_data$Price_MEG <- raw_data$y
raw_data$Price_crude_oil_coal <- 0.6 * raw_data$x1 + 0.4 * raw_data$x3
raw_data$Upstream_MEG_Profit <- 0.5 * raw_data$x4 + 0.5 * raw_data$x5
raw_data$MEG_Profit <- 0.5 * raw_data$x6 + 0.5 * raw_data$x7
raw_data$MEG_operating_rate <- raw_data$x8
raw_data$Downstream_Profits <- 1/7 * (raw_data$x10 + raw_data$x11 + raw_data$x12 + raw_data$x13 + raw_data$x14 + raw_data$x15 + raw_data$x16)
raw_data$Downstream_Operating_Rate <- 1/6 * (raw_data$x17 + raw_data$x18 + raw_data$x19 + raw_data$x20 + raw_data$x21 + raw_data$x22)/100
raw_data$Downstream_Inventory <- 0.25 * (raw_data$x24 + raw_data$x25 + raw_data$x26 + raw_data$x27)
raw_data$Inventory <- 0.25 * raw_data$x28 + 0.5 * raw_data$x29 + 0.25 * raw_data$x30

# Remove old variables and 'y', 'x23'
cols_to_remove <- c("y", "x1", "x2", "x3", "x4", "x5", "x6", "x7", "x8", "x9", "x10", "x11", "x12", "x13", "x14", "x15", "x16", "x17", "x18", "x19", "x20", "x21", "x22", "x23", "x24", "x25", "x26", "x27", "x28", "x29", "x30")
data <- raw_data[, !(names(raw_data) %in% cols_to_remove)]

# View the transformed dataset
data <- na.omit(data)
head(data)
dim(data)

# Apply kernel density estimate for Price_MEG using an Epanechnikov kernel.
bw1 <- npudensbw( ~ Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw1), main = "Kernel density estimate for Price of MEG")

# Conduct Pearson tests for normality.
pearson.test(data$Price_MEG)

# Apply log transformation to the response variable
data$log_Price_MEG <- log(data$Price_MEG)
bw2 <- npudensbw( ~ log_Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw2), main = "Kernel density estimate for Log Price of MEG")

# Apply square root transformation to the response variable
data$sqrt_Price_MEG <- sqrt(data$Price_MEG)
bw3 <- npudensbw( ~ sqrt_Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw3), main = "Kernel density estimate for Square Root Price of MEG")

# Apply inverse transformation to the response variable
data$inv_Price_MEG <- 1/data$Price_MEG
bw4 <- npudensbw( ~ inv_Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw4), main = "Kernel density estimate for Inverse Price of MEG")

# Apply inverse log transformation to the response variable
data$inv_log_Price_MEG <- 1/data$log_Price_MEG
bw5 <- npudensbw( ~ inv_log_Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw5), main = "Kernel density estimate for Inverse Log Price of MEG")

# Drop the transformed response variables
cols_to_remove <- c("log_Price_MEG", "sqrt_Price_MEG", "inv_Price_MEG", "inv_log_Price_MEG")
data <- data[, !(names(data) %in% cols_to_remove)]


```


# Dataset Introduction and Hypotheses
The dataset we are using examines how the price of Mono-ethylene glycol (MEG) is related to upstream and downstream variables. Originally, there are 29 explanatory variables with a lot of missing data. With the help of the domain knowledge, We transformed them into 8 variables using the weighted method. We will be using the new transformed variables in our analysis, including the price of crude oil and coal, upstream MEG profit, MEG profits, MEG operating rate, downstream profits, downstream operating rate, downstream inventory, and inventory. The response variable is the price of MEG. The purpose of this report is to examine how the upstream and downstream variables affect the price of MEG, and to predict the price based on the variables.

The hypotheses are as follows:
We expect that the price of MEG will be positively related to the price of crude oil and coal, the upstream MEG profit, MEG profit, downstream profits, and downstream operating rate. As the price of crude oil and coal is a measure of the cost of MEG, and the price will be higher when the cost is higher. Besides, the upstream MEG profit, MEG profit, downstream profits, and downstream operating rate are all measures of the demand of MEG, and the price will be higher when the demand is higher.

We also expect that the price of MEG will be negatively related to the inventory, as the inventory is a measure of the supply of MEG, and the price will be lower when the supply is higher.

We will also try to find an appropriate model to predict the price of MEG based on the variables.


# Description of the Dataset
```{r, echo=FALSE}
# Create a data frame with the information
df <- data.frame(
  Variable_Name = c("date", "Price_MEG", "Price_crude_oil_coal", "Upstream_MEG_Profit", "MEG_Profit", "MEG_operating_rate", "Downstream_Profits", "Downstream_Operating_Rate", "Downstream_Inventory", "Inventory"),
  Unit_of_Measurement = c("YYYY/M/D", "CNY()/ton", "USD()/barrel", "CNY()/ton", "CNY()/ton", "Percentage", "CNY()/ton", "Percentage", "10,000 Tons", "10,000 Tons"),
  Continuous_vs_Discrete = c("Discrete", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous", "Continuous")
)

# Print the table using kable
#kable(df, caption = "Table 1: Description of the variables")
```
| Variable Name             | Unit of Measurement | Continuous vs Discrete |
|---------------------------|---------------------|------------------------|
| date                      | YYYY/M/D            | Discrete               |
| Price_MEG                 | CNY(짜)/ton          | Continuous             |
| Price_crude_oil_coal      | USD($)/barrel       | Continuous             |
| Upstream_MEG_Profit       | CNY(짜)/ton          | Continuous             |
| MEG_Profit                | CNY(짜)/ton          | Continuous             |
| MEG_operating_rate        | Percentage          | Continuous             |
| Downstream_Profits        | CNY(짜)/ton          | Continuous             |
| Downstream_Operating_Rate | Percentage          | Continuous             |
| Downstream_Inventory      | 10,000 Tons         | Continuous             |
| Inventory                 | 10,000 Tons         | Continuous             |
                       Table 1: Description of the variables


As stated from the Table 1 above, the dataset contains 10 variables. The first variable is the date. Price of MEG is the response variable. All the other variables are explanatory variables. The date is a discrete variable, and the rest of the variables are continuous variables. The dataset collected weekly from 2018/1/5 to 2022/11/4, and there are 239 observations in this dataset. The dataset is collected from the Lili's previous employer, and Lili is allowed to use the dataset for this project.



# Regression Analysis
## Normality check for the response variable
We conduct a kernel density estimate for Price of MEG using an Epanechnikov kernel. The Epanechnikov kernel is a nonparametric kernel that is used to estimate the probability density function of a random variable.
```{r, echo = FALSE}
options(np.messages=FALSE)

# Apply kernel density estimate for Price_MEG using an Epanechnikov kernel.
bw1 <- npudensbw( ~ Price_MEG, data = data, ckertype = "epanechnikov", bwmethod = "cv.ml")
plot(npudens(bws = bw1), main = "Kernel density estimate for Price of MEG",
     xlab = "Price of MEG",
     cex.lab = 0.8, cex.main = 0.9, cex.axis = 0.8)
title(sub = "Figure 1: The kernel plot clearly shows the distribution of the\n response variable is not normal.",
      cex.sub = 0.8, line = 5)
```
Figure 1 clearly shows the distribution of the response variable is not normal. We also conduct a Pearson test for normality. With an extremely small p-value, we reject the null hypothesis that the response variable is normally distributed. Therefore, we tried to transform the response variable to make it more normal first.
```{r, echo = FALSE}
# Create individual ggplot objects
log_Price_MEG_plot <- ggplot(data, aes(x = log(Price_MEG))) +
  geom_density() +
  labs(title = "Density of Log Price of MEG",
       x = "Log Price of MEG")

sqrt_Price_MEG_plot <- ggplot(data, aes(x = sqrt(Price_MEG))) +
  geom_density() +
  labs(title = "Density of Square Root Price of MEG",
       x = "Square Root Price of MEG")

inv_Price_MEG_plot <- ggplot(data, aes(x = 1 / Price_MEG)) +
  geom_density() +
  labs(title = "Density of Inverse Price of MEG",
       x = "Inverse Price of MEG")

inv_log_Price_MEG_plot <- ggplot(data, aes(x = 1 / log(Price_MEG))) +
  geom_density() +
  labs(title = "Density of Inverse Log Price of MEG",
       x = "Inverse Log Price of MEG")

# Create the caption with centered alignment
caption <- ggdraw() + draw_label(
  "Figure 2: Density plots for different transformations of Price of MEG.",
   x = 1, hjust = 0.5, vjust = 1, size = 12)

# Arrange the plots in a grid layout with the caption
plot_grid(log_Price_MEG_plot, sqrt_Price_MEG_plot, inv_Price_MEG_plot, inv_log_Price_MEG_plot, caption,
          ncol = 2, nrow = 3, rel_heights = c(1, 1, 0.1))
```


As shown in Figure 2, after using log, square root, inverse transformation, and inverse log transformation, we found that all of them can not handle the skewness and multi-peaks of the response variable. So we decided that nonparametric regression is the best choice for this dataset.

## Check the correlation between variables
Before fitting a nonparametric regression model, we need to check the correlation between the response variable and the explanatory variables.


```{r, echo = FALSE}
# Check correlation between variables and plot the correlation plot with coefficients displayed
corrplot(cor(data[, -1]), method = "color", type = "upper",
         addCoef.col = "black", # Color of the correlation coefficients
         number.cex = 0.8,      # Font size of the correlation coefficients
         tl.col = "black",      # Color of the variable names
         tl.srt = 45,           # Text rotation angle for variable names
         tl.cex = 0.8,          # Font size of the variable names
         diag = FALSE)          # Do not display diagonal elements
# Add a caption below the plot
mtext("Figure 3: Correlation plot with coefficients between variables",
      side = 1, line = 4, cex = 1.1)

```

Figure 3 shows that the price of MEG is positively highly correlated with MEG profit and inventory, and has almost no correlation with the price of crude oil and coal. Surprisingly, we expected that the price of MEG will be positively correlated with the price of crude oil and coal, and negatively correlated with the inventory.

## Nonparametric local linear regression
```{r, include = FALSE}
options(np.messages=FALSE)

set.seed(8888)
bw <- npregbw(Price_MEG ~ Price_crude_oil_coal + Upstream_MEG_Profit + MEG_Profit + MEG_operating_rate + Downstream_Profits + Downstream_Operating_Rate + Downstream_Inventory + Inventory,
              data = data, regtype = "ll", bwmethod = "cv.aic")
np_local_lm <- npreg(bws = bw)
summary(np_local_lm)

result <- npsigtest(np_local_lm)
```
After fitting the nonparametric local linear regression model, the summary of the model shows the R-squared is 0.997, which raises the concern of over-fitting. We then performed the consistent nonparametric test of significance, we found that the p-value of the price of crude oil and coal, upstream MEG profit, MEG profit, MEG operating rate are smaller than 0.05, which means they are significant. All the other variables are not significant according to the test.


## Generalized additive model
```{r, echo = FALSE, message = FALSE, warning = FALSE}
gam_model <- gam(Price_MEG ~ s(Price_crude_oil_coal) + s(MEG_Profit),
                 data = data, family = gaussian)
#summary(gam_model)

par(mfrow=c(1,2))
# Plot for Price_crude_oil_coal
plot(gam_model, select = 1, xlab = "Price crude oil coal", ylab = "Partial effect")
# Plot for MEG_Profit
plot(gam_model, select = 2, xlab = "MEG Profit", ylab = "Partial effect")

title(sub = "Figure 4: Marginal splines for Price crude oil coal and MEG Profit",
      cex.sub = 1, line = 4)


# # Construct a 3D plot
# x1 <- seq(from = 200, to = 1000, by = 10)
# x2 <- seq(from = -2000, to = 3000, by = 50)
# x_grid <- data.frame(expand.grid(Price_crude_oil_coal = x1, MEG_Profit = x2))
#
# gam_pred <- predict(gam_model ,newdata = x_grid)
# surface <-  matrix(gam_pred, nrow=length(x1))
#
# open3d()
# persp3d(x = x1, y = x2,
#         z = surface, theta = 30, phi = 30, expand = 0.5,
#         col = "orange", xlab = "Price crude oil coal",
#         ylab = "MEG Profit", zlab = "Predicted Price MEG")
# points3d(data$Price_crude_oil_coal ~ data$MEG_Profit + data$Price_MEG, col="blue")
```

We fitted a generalized additive model with the Price of MEG as the response variable and the Price of crude oil and coal and MEG profit as the explanatory variables. The summary of the model shows that the adjusted R-squared is 0.945, which is similar to the nonparametric local linear regression model.

This partial effect plots in Figure 4 can show the relationships between the response variable and the explanatory variable, while accounting for the effect of the other explanatory variable.

Both plots show that the Price of MEG increases as the Price of crude oil and coal, and MEG profit increase. The right plot shows the relationship between the Price of MEG and the MEG profit is more linear compared to the left plot. The left plot also has more uncertainty in the fitted smooth function, which means that the relationship between the Price of MEG and the Price of crude oil and coal is more complex.

## Variable importance
```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Random forest
rf <- randomForest(Price_MEG ~ ., data = data[,-1], importance = T)
varImpPlot(rf)
title(sub = "Figure 5: Variable importance plot for random forest",
      cex.sub = 1, line = 4)

boosting <-  gbm(Price_MEG ~ ., data = data[,-1], distribution = "gaussian",
                 n.trees = 10000, interaction.depth = 6, shrinkage = 0.03, bag.fraction = 0.8)
summary(boosting)
```
We check the variable importance using random forest and boosting. Both results show that inventory and MEG profit are the 2 most important variables, followed by the Price of crude oil and coal. This results agree with the nonparametric local linear regression model that MEG profit and price of crude oil and coal are important variables, but inventory is not significant in the nonparametric local linear regression model.

Overall, for the purpose of variable selection, if we would like to keep 2 variables, we would choose MEG profit and inventory. If we would like to keep 3 variables, we would choose MEG profit, inventory, and the price of crude oil and coal.

##

##


# Conclusion


