---
title: "Exploratory analysis"
output:
  pdf_document: default
  html_document: default
---

```{r, include = FALSE}
library(ggplot2)
library(gridExtra)
library(glmnet)
library(np)
library(nortest)
library(MASS)
library(glmnet)
library(leaps)
```
```{r}
# getwd()
#setwd('DATA 583/583-project')

# Load the data from the CSV file
data <- read.csv("data_cleaned.csv", header = TRUE, check.names = FALSE)
dim(data) # 189, 31

#remove date
data_no_date <- data[, -1]

head(data)
```

We aim to predict the price of ethylene glycol using the following 29 explanatory variables.
MEG refers to ethylene glycol.
y	MEG spot price

Up-stream price of MEG
x1	WTI: West Texas Intermediate（Crude Oil）Price
x2	Brent: Brent Crude oil price
x3	Coal price

Up-stream(Ethylene) Profit
x4	domestic
x5	foreign

MEG Profit
x6	made of coal
x7	made of Ethylene

MEG operating rate
x8	domestic
x9	foreign

Downstream profits
x10	Recycled Bottle Chips profits
x11	polyester chips profits
x12	polyester bottle chip profits
x13	POY profits
x14	FDY profits
x15	DTY profits
x16	Polyester profits

Downstream operating rate
x17	Polyester
x18	filament
x19	Direct spinning
x20	chip spun filament
x21	texturing machine
x22	weaving machine
x23	Dyeing Factory

Downstream Inventory
x24	Polyester
x25	FDY
x26	DTY
x27	POY

x28	MEG foreign Port Inventory
x29	MEG domestic Port Inventory
x30	MEG factory Inventory

The data structure is shown below. We can see that all the variables are continuous values.

```{r}
# The structure of the data,
str(data)
```

Here is the summary statistics of the data.
```{r}
summary(data)
```

Let's explore the histogram of the response variable y and some of the explanatory variables.

```{r, warning = FALSE, message = FALSE}
# Create histograms of some of the variables
hist_y <- ggplot(data, aes(x = y)) + geom_histogram() + labs(title = "Histogram of y")
hist_x1 <- ggplot(data, aes(x = x1)) + geom_histogram() + labs(title = "Histogram of x1")
hist_x11 <- ggplot(data, aes(x = x11)) + geom_histogram() + labs(title = "Histogram of x11")
hist_x21 <- ggplot(data, aes(x = x21)) + geom_histogram() + labs(title = "Histogram of x21")

grid.arrange(hist_y, hist_x1, hist_x11, hist_x21, nrow = 2, ncol = 2)
```

The histogram of the response variable y seems to have 3 peaks. The histogram of the explanatory variable x1 and x11 are right skewed. The histogram of the explanatory variable x21 is left skewed.


Let's explore the scatterplots of the response variable y and some of the explanatory variables.
```{r}
# create scatterplots of the variables x1, x2, x3, x4, x5, x7, x10, x11, x12
scatter_x1 <- ggplot(data = data, aes(x = x1, y = y)) + geom_point() + labs(title = "Scatterplot of x1 and y")
scatter_x7 <- ggplot(data = data, aes(x = x7, y = y)) + geom_point() + labs(title = "Scatterplot of x7 and y")
scatter_x14 <- ggplot(data = data, aes(x = x14, y = y)) + geom_point() + labs(title = "Scatterplot of x14 and y")
scatter_x30 <- ggplot(data = data, aes(x = x30, y = y)) + geom_point() + labs(title = "Scatterplot of x30 and y")

grid.arrange(scatter_x1, scatter_x7, scatter_x14, scatter_x30, ncol = 2, nrow = 2)
```
There seems to have a positive linear relationship between the response variable y and the explanatory variables x1, x7. And a negative relationship between y and x30. There is no obvious relationship between y and x14.

Let's explore the trend of the response variable y and some of the explanatory variables over time.

```{r}
# plot the variables versus date
trend_y <- ggplot(data = data, aes(x = date, y = y)) + geom_point() + labs(title = "Trend of y over time")
trend_x1 <- ggplot(data = data, aes(x = date, y = x1)) + geom_point() + labs(title = "Trend of x1 over time")
trend_x4 <- ggplot(data = data, aes(x = date, y = x4)) + geom_point() + labs(title = "Trend of x4 over time")
trend_x11 <- ggplot(data = data, aes(x = date, y = x11)) + geom_point() + labs(title = "Trend of x11 over time")

grid.arrange(trend_y, trend_x1, trend_x4, trend_x11, ncol = 2, nrow = 2)
```

The trend plots of y and x1 seem to have some similarities. The trend plots of x4 and x11 have no obvious trend.

Let's try to plot a kernel density estimate for y using an Epanechnikov kernel.
```{r}
bw <- npudensbw( ~ y, data = data,
                 ckertype = "epanechnikov", bwmethod = "cv.ml")

fhat <- npudens(bws = bw)
plot(fhat)
```
As stated in the histogram part, the distribution of y is not normal and it seems to have 3 peaks.

Let's conduct Pearson tests for normality.
```{r}
pearson.test(data$y)
pearson.test(data$x11)
```
From the pearson chi-square normality tests, we can see that the p-value is very small for y, and large for x11.
So we have evidence that y is not normally distributed, but x11 is normally distributed.


Let's try a linear regression model to do diagnostics.
```{r}
# fit a linear regression model
model_ls <- lm(y ~ ., data = data_no_date)
summary(model_ls)

plot(model_ls)
```
From the summary output, the can see that only variables x3, x6, x7, x13, x14, x20, x22, x25, x27, x28 and x30 are significant.
The R-squared and adjusted R-square are high, which is expected as we have lots of variables. Variable selection is needed to reduce the number of variables.

Let's try to do variable selection using stepwise regression.
```{r}
model_step <- stepAIC(model_ls, direction = "both", criterion = "bic", trace = FALSE)
model_step
```
From the stepwise model selection, 6 variables are removed from the model.

Let's try to do variable selection using lasso regression.
```{r}
# fit a lasso regression model
model_lasso <- cv.glmnet(x = as.matrix(data_no_date[, -1]), y = data_no_date[, 1], alpha = 1)

# plot the lasso regression model
plot(model_lasso)
```
Both minimum and 1se lines suggest to keep 23 variables, which is the same as the stepwise model selection.









